---
title: "PCA on Regional Data"
author: "Corey Pang"
date: "7/12/2017"
output: html_document
---

```{r setup, include=FALSE}
install.packages("devtools", repos = "http://cran.us.r-project.org")
library(devtools)
install_github("vqv/ggbiplot", force=TRUE)
library(ggbiplot)
getwd()
dir <- file.path(getwd(),"Regional data")
out <- read.csv(file.path(dir, "LandCoverPercentsByDistrict.csv"))
is.na(out)
```

#### PCA on LULC Variables 
```{r, incluse=FALSE}
reg.data <- out[,2:11]
reg.pca <- prcomp(reg.data, scale = TRUE, center = TRUE)
summary(reg.pca)
sd <- reg.pca$sdev
loadings <- reg.pca$rotation
rownames(loadings) <- colnames(reg.data)
scores <- reg.pca$x
```

```{r1}
var <- sd^2
var.percent <- var/sum(var) * 100
barplot(var.percent, xlab="PC", 
        ylab="Percent Variance", 
        main = "Barplot of Percentage of Total Variation",
        names.arg=1:length(var.percent), 
        las=1, ylim=c(0,max(var.percent)), 
        col="gray")
abline(h=1/ncol(reg.data)*100, col="red")
var.percent[1:4]
sum(var.percent[1:4])
```
There is a big drop in percent variance explained from PC1 to PC2. As we can see from the barplot, 4 principal components explain more than one variable’s worth of information so we will look into these principal components. These 4 components explain 71% of the variation in the data. 

```{r2}
plot(var.percent, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             main = "Scree Plot for LULC Variables",
             type = "b")
```

Cutoff for Important Loadings
```{r3}
sqrt(1/ncol(reg.data))
reg.loadings <- reg.pca$rotation[,1:4]
print(reg.loadings, digits = 2, cutoff = sqrt(1/ncol(water)), sort = TRUE)
```

The percent of Wetlands and Salt have a moderately positive association with PC1. Cropland has a slightly lower positive association with PC1 than Wetlands and Salt. Urban and Sand barely make the cut off for important loadings, but they have a weak positive association with PC1. Percent Riparian, Urban, and Cloud both have a weakly moderate association with PC2. Sand percentage has a weak negative association with PC2. For PC3, Shrubland has a relatively strong positive association with PC3 and and relatively weak association with Forest. Water percent has a strong negative association with PC4 and a weak to moderate negative association with Riparian. Cropland has a positive weak association with PC4. 



```{r4}
print(ggbiplot(reg.pca, scale = 1, obs.scale = TRUE, var.scale = TRUE))
```




