---
title: "PCA on Regional Data"
author: "Corey Pang"
date: "7/12/2017"
output: html_document
---

```{r setup, include=FALSE}
install.packages("devtools", repos = "http://cran.us.r-project.org")
library(devtools)
install_github("vqv/ggbiplot", force=TRUE)
library(ggbiplot)
library(foreign)
library(shapefiles)
library(reshape2)
library(psych)
getwd()
dir <- file.path(getwd(),"Regional data")
out <- read.csv(file.path(dir, "LandCoverPercentsByDistrict.csv"))
populations <- read.csv(file.path(dir,"finalDistrictPop.csv"))
healthCenter = read.dbf(file.path(dir,"healthCenterCountsByDistrict.dbf"))
healthCenterup = data.frame(ADM2 = healthCenter$dbf$ADM2, hcCount= healthCenter$dbf$Count_)
hospital = read.dbf(file.path(dir,"HospitalCountsByDistrict.dbf"))
hospitalup = data.frame(ADM2 = hospital$dbf$ADM2, hospCount= hospital$dbf$Count_)
dataHealth = merge(healthCenterup, hospitalup, by = 'ADM2')
updated = merge(dataHealth, out, by='ADM2')
updated = merge(updated, populations, by = 'ADM2')
```

#### PCA on LULC Variables 
```{r, incluse=FALSE}
reg.data <- out[,2:11]
reg.pca <- prcomp(reg.data, scale = TRUE, center = TRUE)
summary(reg.pca)
sd <- reg.pca$sdev
loadings <- reg.pca$rotation
fa.sort(loadings)
rownames(loadings) <- colnames(reg.data)
scores <- reg.pca$x
```

```{r1}
var <- sd^2
var.percent <- var/sum(var) * 100
barplot(var.percent, xlab="PC", 
        ylab="Percent Variance", 
        main = "Barplot of Percentage of Total Variation",
        names.arg=1:length(var.percent), 
        las=1, ylim=c(0,max(var.percent)), 
        col="gray")
abline(h=1/ncol(reg.data)*100, col="red")
var.percent[1:4]
sum(var.percent[1:4])
```
There is a big drop in percent variance explained from PC1 to PC2. As we can see from the barplot, 4 principal components explain more than one variable’s worth of information so we will look into these principal components. These 4 components explain 71% of the variation in the data. 

```{r2}
plot(var.percent, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             main = "Scree Plot for LULC Variables",
             type = "b")
```

Cutoff for Important Loadings
```{r3}
sqrt(1/ncol(reg.data))
reg.loadings <- reg.pca$rotation[,1:4]
print(reg.loadings, digits = 2, cutoff = sqrt(1/ncol(water)), sort = TRUE)
```

The percent of Wetlands and Salt have a moderately positive association with PC1. Cropland has a slightly lower positive association with PC1 than Wetlands and Salt. Urban and Sand barely make the cut off for important loadings, but they have a weak positive association with PC1. Percent Riparian, Urban, and Cloud both have a weakly moderate association with PC2. Sand percentage has a weak negative association with PC2. For PC3, Shrubland has a relatively strong positive association with PC3 and and relatively weak association with Forest. Water percent has a strong negative association with PC4 and a weak to moderate negative association with Riparian. Cropland has a positive weak association with PC4. 



```{r4}
# Changed data into one column of values to use for ggplot2

loadings.m <- melt(loadings, id="ID", 
                   measure=c("PC1", "PC2", 
                             "PC3", "PC4"), 
                   variable.name="Factor", value.name="Loading")
ggplot(loadings.m, aes(x = ID, y = Loading, fill=Loading)) + 
  facet_wrap(~ Factor, nrow = 1) +
  geom_bar(stat="identity")
```

str(reg.loadings)
summary(loadings)
id <- seq(1,10,1)
cbind(loadings, id)
Ord <- c(1, 2, 4, 3, 5, 6, 7, 8, 10, 9)
loadings$id <- reorder(loadings$id, Ord) 
loadings <- data.frame(ID = rownames(reg.pca$rotation), reg.pca$rotation[,1:4], stringsAsFactors=F)
loadings <- loadings[order(loadings$PC1, decreasing = F),]

ggplot(loadings$PC1, aes(x = ID, y = PC1, decreasing=T))
